This project deals with modeling and control of a refrigeration trailer. Refrigeration trailers are important for transportation of perishable goods such as fresh meat and fish. The refrigeration trailer maintains appropriate temperature levels to preserve the durability of the cargo.
This project is conducted in cooperation with BITZER Electronics A/S as a semester project at 2nd semester of the Masters Programme in Control and Automation at Aalborg University.

\begin{enumerate}
	\item Why the need for reefer units in general?
	\item Why the need for stable and accurate cooling?: Goods with strict requirements.
	% Relevant literature: https://www.tis-gdv.de/tis_e/containe/arten/kuehlcon/kuehlcon.htm/
	\item Why the need for efficiency?: Battery - Drive longer or use smaller and cheaper battery
	\item How reefer trailer work generally: They \underline{maintain} temperature of goods
	\item The main disturbances in reefer trailers.
\end{enumerate}

This project deals with modeling and control of a refrigeration trailer whose purpose is to maintain specific temperatures inside a box (I know what you're thinking.. Using "whose" for things is grammatically incorrect. To this i say: Shakespeare used it an thus i will as well!!!). The necessity for the transportation of perishable goods such as fresh meat and fish can hardly be questioned. All sorts of products require a temperature-controlled environment which will depend and vary based on the specific product. Some examples are frozen food, dairy products, produce, medical supplies (including vaccines), and electronics. This further highlights the need for high quality temperature-controlled transportation options. While reefer \textit{containers} are used in shipping, reefer \textit{trailers} are necessary when transporting to warehouses and grocery stores.

Unlike reefer containers, reefer trailers utilize a battery as the main power supply due to the limited power delivery of the trucks they are hooked upon. Efficiency is therefore of great importance as this allows for not only greater transportation distances but also a possible reduction of the battery pack size which is appealing because of their high cost.

The purpose of a reefer trailer is to maintain the temperature of its cargo and not actually cooling it down. It is thus expected that cargo is pre-cooled to around the desired transportation temperature before being loaded. A desired temperature is reached by cooling down air which is circulated between the cooling box and the refrigeration unit. The refrigeration unit is situated at the back of the trailer where all components besides sensors are located. Inside the refrigeration unit a refrigerant is moved between a low pressure in the evaporator and high pressure in the condenser. The evaporator subtracts heat from the circulating box air and the condenser gives the heat to the air outside the box.